---
title: "Homework 1"
author: "[YOUR NAME HERE]"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    highlight: textmate
    theme: readable
    toc: yes
    toc_float: yes
---

<!-- Feel free to edit theme / highlight in the YAML header above -->

```{r setup, echo=TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = TRUE)

# Load libraries for homework problems
library(gapminder) # for data
library(ggplot2)   # for graphs
library(readr)     # for reading in solutions
library(gganimate) # for animation
library(gifski)    # for animation


```

# Problem 1

We will be using a curated version of the `gapminder` dataset. 

1. Create a logical vector using `gapminder` that is `TRUE` if `gapminder$year` is equal to 2007 and `FALSE` otherwise. 

1. Use the vector to create a subset of the gapminder data, containing only the rows from the year 2007. Call this subset of data `gapminder_2007`

**Note**: If you get stuck on this problem, you can proceed to the next problems using the `gapminder_2007` object produced by the solution code below. 

Your 2007 data should look like this:

```{r}

gapminder_2007 <- read_rds('solutions/01_solution.rds')

gapminder_2007

```

# Problem 2

Create a plot using the `ggplot2` package:

- use the `ggplot()` function and specify `data = gapminder_2007`.

- use the `aes()` function:

    + y axis should be `lifeExp` (life expectancy)

    + x axis should be `gdpPercap` (gross domestic product per capita; a measure of income)
    
    + the `fill` of objects plotted should correspond with `continent` values.
    
    + the `size` of objects plotted should correspond with `pop` values.

- use `geom_point(shape = 21)` to get points with visible fill colors.

- adjust the x-axis to be on the log 10 scale. 

The finished figure will look like this:

```{r}

read_rds('solutions/02_solution.rds')

```  

# Problem 3

Make some aesthetic changes to the figure.

- add labels to the fill, scale, title, subtitle, and axes of the figure.

- use a `theme_()` function of your choice. (mine is `theme_minimal()`)

- use the `theme()` function to increase the size of all text in the plot.

The finished figure will look like this:

```{r}

read_rds('solutions/03_solution.rds')

```

# Problem 4

Make a couple more adjustments to the figure:

- Remove the point size guideline from the legend.

- Increase the size of points in the fill portion of the legend. (hint: look up `guides`)

```{r}

read_rds('solutions/04_solution.rds')

```

# Problem 5

- Re-create the figure from problem 4 using the entire `gapminder` dataset instead of the `gapminder_2007` data.

- Use the `gganimate` package to make a smooth animation that has frames based on the `year` column. 

*Note:* You can find a great `gganimate` tutorial [here](https://github.com/thomasp85/gganimate)

```{r, cache = TRUE}

soln <- read_rds('solutions/05_solution.rds')

animate(soln,
  nframes = 250,
  fps = 50,
  width = 96 * 8,
  height = 96 * 6
)

```

